---
title: "Intro to Linear Regression"
author: "Jessica Madden"
date: "November 19, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(ggpubr)
library(broom)
```

# A rank-based test exmple: Mann Whitney U

```{r}
set.seed(1414) #creates a pseudorandom sample
gp_1 <- sample.int(20, size = 15, replace = TRUE) #creates random samples with integers 1-x
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

### Exploratory histogram
Is there evidence for a significant difference in ranks (medians) between the populations from which gp1 and gp2 are drawn?

```{r}
hist(gp_1)
```

```{r}
hist(gp_2)
```

### Mann Whitney U 

```{r}
my_mwu <- wilcox.test(gp_1, gp_2)
```


# Simple Linear Regression

### A. Look at it!

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

### B. Model it

```{r}
#Linear model, stored as penguin_lm:
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

#Return the somplete overview:
summary(penguin_lm)
```

### C. Access model outputs

use the broom::tidy() function to get the model outputs in nice data frame format:

```{r}
penguin_lm_tidy <- broom::tidy(penguin_lm)
```

```{r}
# Get the intercept:
penguin_int <- penguin_lm_tidy$estimate[1]
penguin_int
```

```{r}
# Then get the flipper_length coefficient:
penguin_coef <- penguin_lm_tidy$estimate[2]
penguin_coef
```

```{r}
# Metrics at a glance:
penguin_lm_out <- broom::glance(penguin_lm)
penguin_lm_out
```

Write statement about model that will automatically update:

"Simple linear regression was used to explore the relationship between penguin flipper length (mm) and body mass (g) across all three penguin species, and including both male and female penguins. A significant regression model was found ($\beta$ = `r round(penguin_coef,3)`, F(`r penguin_lm_out$df`,`r penguin_lm_out$df.residual`) = `r round(penguin_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(penguin_lm_out$r.squared,3)`."

### D. Explore model assumptions

```{r}
plot(penguin_lm)
```

###. Visualize the model

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", 
              color = "red", 
              size = 0.5, 
              fill = "gray10", 
              alpha = 0.5) +
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700) #add equation information directly to the plot panel
```

### F. Find Pearson's R for correlation

```{r}
penguins_cor <- cor.test(penguins$flipper_length_mm, penguins$body_mass_g)
penguins_cor
```

Here, we see that there is a strong positive correlation between penguin flipper length and body mass (*r* = `r round(penguins_cor$estimate,2)`, t(`r penguins_cor$parameter`) = `r round(penguins_cor$statistic,2)`, p < 0.001). 